\sffamily An EEG-based visual Brain-Computer Interface for gaze-free communication

### 2024-10-09 chapters/stbf_struct.tex
Comments François patients

  \emph{This chapter was
  published as~\textcite{VanDenKerchove2022}}.

@@

	For the following analyses, we assume that all \ac{eeg} channels are normalized with zero mean and unit variance without loss of generality.
	Solving \cref{eq:minimum_variance} under the linear constraint given by
	\cref{eq:linear_constraint} returns the filter weights $w$ defining the spatiotemporal \ac{lcmv}-beamformer.
	\begin{equation}
    \underset{\mat{w}}{\arg\min}\mat{w}^\intercal \mat{C}
		\mat{w}^\intercal
		\label{eq:minimum_variance}
	\end{equation}
	\begin{equation}
		\mat{a}^\intercal\mat{w} = 1
		\label{eq:linear_constraint}
	\end{equation}

	These weights minimize the variance of the output of the filter while enhancing
	the signal characterized by the constraint.
  $\mat{a} = \text{vec}(\mat{A})$ is the data-driven activation pattern, a template
	of the signal of interest maximizing the difference between two classes of
	epochs, determined as follows:

	\begin{equation}
		\mathbf{a} =
		\frac{1}{|\text{class 1}|}\sum_\text{class 1}\mathbf{x_i} -
		\frac{1}{|\text{class 2}|}\sum_\text{class 2}\mathbf{x_i}
		\label{eq:activation_pattern}
	\end{equation}

	The method of Lagrange multipliers then gives the closed-form solution to the minimization problem posed by
	\cref{eq:minimum_variance} and \cref{eq:linear_constraint} as:

	\begin{equation}
		\mat{w} =
    \frac{\mat{C}^{-1}\mat{a}^\intercal}
		{\mat{a}C^{-1}\mat{a}^\intercal}
		\label{eq:closed_form}
	\end{equation}

	The beamformer can be applied to epochs (unseen or not) as:

	\begin{equation}
		y_i = \mathbf{w}\mathbf{x}_i
		\label{eq:apply_beamformer}
	\end{equation}

	resulting in a scalar output per epoch.
	The linear constraint in \cref{eq:linear_constraint} ensures that the
	beamformer maps epochs containing a target response to a score close to one
	and, conversely, epochs not containing a target response to a score close to
	zero.

	\subsection{Covariance matrix regularization}
	While the spatiotemporal beamformer, in theory, achieves optimal separation
	between target and non-target classes, in analogy to linear discriminant
  analysis~\cite{Treder201}, it does not always perform well on unseen data.
	The main challenge is to find a good estimator for the inverse covariance
  matrix $\mat{C}^{-1}$ since the real underlying covariance matrix generating the data is, in principle, unknown.

	\subsubsection{Empirical covariance estimation}
	\label{sec:empirical_covariance}
	Earlier spatiotemporal beamformer studies~\cite{Wittevrongel2016,
		Wittevrongel2016a, Wittevrongel2017, Wittevrongel2017a} use the empirical
	covariance and inverse covariance calculated as follows:

	\begin{equation}
    \hat{\mat{C}}_\text{emp} =
		\frac{1}{N-1}\sum^{N}_{i=1}\mat{x}_i\mat{x}_i^\intercal
		\label{eq:empirical_covariance}
	\end{equation}

	\begin{equation}
    \widehat{\mat{C}^{-1}}_\text{emp} = \hat{\mat{C}}_\text{emp}^+
		\label{eq:empirical_inverse_covariance}
	\end{equation}

	The Moore-Penrose pseudoinverse $^+$, ensures a solution exists when
  $\hat{\mat{C}}_\text{emp}$ is singular.
	\cref{fig:kronlda-covs}a and \cref{fig:kronlda-covs}b respectively show examples of the
	empirical estimators of the covariance and the inverse covariance matrices.
	The empirical estimator suffers from performance and
	stability issues if the number of epochs $n$ used or estimation is not much larger than the number of features $cs$~\cite{Stein1956,Khatri1987}.

	\subsubsection{Shrunk covariance estimation}
	\label{sec:shrunk_covariance}
	The shrinkage covariance estimator creates a better-conditioned inversion matrix problem and generally performs better when applied to unseen data.
	The estimators for the covariance and inverse covariance are given by:

	\begin{equation}
    \hat{\mat{C}}_\alpha =
    (1-\alpha) \hat{\mat{C}}_\text{emp}
    + \alpha\frac{\text{Tr}(\hat{\mat{C}}_\text{emp})}{CS}\mathbb{I}
		\label{eq:shrinkage}
	\end{equation}

	\begin{equation}
    \widehat{\mat{C}^{-1}}_\alpha =
    \hat{\mat{C}}^+_\alpha
		\label{eq:shrinkage_inverse}
	\end{equation}

	with $0<\alpha<1$.
	Analogous to $L_2$ regularization of the beamforming problem,
	shrinkage reduces the ratio between the smallest and largest eigenvalues
	of the covariance matrix by strengthening the diagonal.
	\cref{fig:kronlda-covs}c and \cref{fig:kronlda-covs}d respectively show examples of the
	shrunk estimator of the covariance and the inverse covariance matrices.

	Earlier work~\cite{Libert2021} applied shrinkage regularization to \ac{erp}
	decoding with the spatiotemporal beamformer and showed competitive performance
	compared to other state-of-the-art decoding techniques like stepwise LDA or SVM.
	The abovementioned work chooses the shrinkage coefficient $\alpha$ as a fixed hyperparameter.
	However, its optimal value depends on the number of training epochs, the
	covariance matrix's dimensionality, and the independence and variance of the
	data, which can vary across evaluation settings and per session.
	The optimal value for $\alpha$ can be found with a line search using cross-validation, but this can be a costly procedure.
	Methods exist to estimate an optimal shrinkage value from the data directly.
	Most notable among these are the Ledoit-Wolf procedure~\cite{Ledoit2004},
	Rao-Blackwell Ledoit-Wolf~\cite{Chen2010}, and Oracle Approximating Shrinkage~\cite{Chen2010}.
	A more recent estimation method~\cite{Tong2018} emulates a leave-one-out
	cross-validation (\ac{loocv}) scheme expressed by the data-driven closed-form
	estimate:

	\begin{equation}
		\alpha =
		1-\frac{
      \frac{N}{N-1}\text{Tr}(\hat{\mat{C}}_\text{emp}^2)
      - \frac{2}{CS}\left[\text{Tr}(\hat{\mat{C}}_\text{emp})\right]^2
      + \frac{1}{CS}\text{Tr}(\hat{\mat{C}}_\text{emp}^2)
			- \frac{1}{N(N-1)}\sum_{i=1}^N||\mat{x}_i||_2^4
		}
		{
      \frac{N^2 -2N}{(N-1)^2}\text{Tr}(\hat{\mat{C}}_\text{emp}^2)
      - \frac{2}{CS}\left[\text{Tr}(\hat{\mat{C}}_\text{emp})\right]^2
      + \frac{1}{CS}\text{Tr}(\hat{\mat{C}}_\text{emp}^2)
			+ \frac{1}{N(N-1)^2}\sum_{i=1}^N||\mat{x}_i||_2^4
		}
		\label{eq:loocv}
	\end{equation}

@@

	\subsubsection{Spatiotemporal beamforming with Kronecker-Toeplitz structured covariance}
	\label{seq:structured_estimation}
	Exploiting prior knowledge about the spatiotemporal structure of the \ac{eeg} signal leads to a more regularized estimator of the covariance.
	When viewing the example of empirical spatiotemporal \ac{eeg} covariance in
	\cref{fig:kronlda-covs}a, it becomes clear that this matrix consists of a block pattern of repeated, similar matrices.
	Due to the multi-channel nature of the signal, we assume that the covariance of spatiotemporal \ac{eeg} epochs is a Kronecker
	product of two smaller
	matrices~\cite{Munck1992,DeMunck1999,Huizenga2002}, as expressed
	by:

	\begin{equation}
    \hat{\mat{C}}_\text{struct} = \hat{\mat{S}} \otimes \hat{\mat{T}}
		\label{eq:kronecker}
	\end{equation}

	with $\otimes$ the Kronecker product operator.
  $\hat{\mat{S}} \in \mathbb{R}^{C\times C}$ and $\hat{\mat{T}} \in \mathbb{R}^{S\times S}$ respectively correspond to estimators of the spatial and temporal covariance of the data.
	Furthermore, because the temporal covariance of the \ac{eeg}-signal is
	stationary (i.e., it is only dependent on interval length between covarying
	time samples)~\cite{Bijma2003}, it is assumed to have a Toeplitz-matrix structure:

	\begin{equation}
    \hat{\mat{T}}_{i,j} = \hat{\mat{T}}_{i+1,j+1}
		\label{eq:toeplitz}
	\end{equation}

	\Cref{prop:inverse_kronecker} then leads to
	\cref{eq:cov_inverse_kronecker} to estimate the inverse
	covariance.
	\begin{property}
    $(\mat{U} \otimes \mat{V})^+ = \mat{U}^+ \otimes \mat{V}^+$ for any non-singular
    matrices $\mat{U}$ and $\mat{V}$~\cite{Langville2004}.
		\label{prop:inverse_kronecker}
	\end{property}

@@

	Using \cref{eq:closed_form_kronecker} removes the need to calculate the
  full, high dimensional Kronecker product $\hat{\mat{S}}^+\otimes
    \hat{\mat{T}}^+$.
	\Cref{fig:kronlda-covs}e and \cref{fig:covs}f respectively show examples of the
	structured covariance and inverse covariance estimators,
	consisting of a spatial Kronecker factor (\cref{fig:kronlda-covs}g and
	\cref{fig:kronlda-covs}h) and a temporal component (\cref{fig:kronlda-covs}i and
	\cref{fig:kronlda-covs}j).

@@

	\begin{subequations}
		\begin{equation}
			\beta_{k+1} =
			1-
			\frac{
        \frac{N}{N-1}\text{Tr}(\tilde{\hat{S}}_{k+1}^2)
        - \frac{2}{C}\left[\text{Tr}(\tilde{\mat{S}}_{k+1})\right]^2
        + \frac{1}{C}\text{Tr}(\tilde{\mat{S}}_{k+1}^2)
				- \frac{1}{N(N-1)}\sum_{i=1}^N
        \left[\text{Tr}(\mat{X}_i\hat{\mat{T}}_k^+\mat{X}_i^\intercal)^2\right]
			}{
        \frac{N^2-2N}{(N-1)^2}\text{Tr}(\tilde{\mat{S}}_{k+1}^2)
        - \frac{2}{C}\left[\text{Tr}(\tilde{\mat{S}}_{k+1})\right]^2
        + \frac{1}{C}\text{Tr}(\tilde{\mat{S}}_{k+1}^2)
				+ \frac{1}{N(N-1)^2}\sum_{i=1}^N
        \left[\text{Tr}(\mat{X}_i\hat{\mat{T}}_k^+\mat{X}_i^\intercal)^2\right]
			}
			\label{eq:spatial_shrinkage}
		\end{equation}
		\begin{equation}
			\gamma_{k+1} =
			1-
			\frac{
        \frac{N}{N-1}\text{Tr}(\tilde{\mat{T}}_{k+1}^2)
        - \frac{2}{S}\left[\text{Tr}(\tilde{\mat{T}}_{k+1})\right]^2
				+ \frac{1}{S}\text{Tr}(\tilde{T}_{k+1}^2)
				- \frac{1}{N(N-1)}\sum_{i=1}^N
        \left[\text{Tr}(\mat{X}_i^\intercal\mat{\mat{S}}_k^+\mat{X}_i)^2\right]
			}{
        \frac{N^2-2N}{(N-1)^2}\text{Tr}(\tilde{\mat{T}}_{k+1}^2)
        - \frac{2}{S}\left[\text{Tr}(\tilde{\mat{T}}_{k+1})\right]^2
        + \frac{1}{S}\text{Tr}(\tilde{\mat{T}}_{k+1}^2)
				+ \frac{1}{N(N-1)^2}\sum_{i=1}^N
        \left[\text{Tr}(\mat{X}_i^\intercal\hat{\mat{S}}_k^+\mat{X}_i)^2\right]
			}
			\label{eq:temporal_shrinkage}
		\end{equation}
	\end{subequations}

	The shrinkage parameters $0<\beta_{k+1}<1$ and $0<\gamma_{k+1}<1$ should be
	re-determined after each iteration.
	The Oracle Approximation Shrinkage method can also be used to determine
	$\beta_{k+1}$ and $\gamma_{k+1}$~\cite{Chen2010,Xie2021} but performs worse for spatiotemporal \ac{eeg} data since not all assumptions are met.

@@

  \subsubsection{\Acl{stbf} classifier}
	Before calculating the \ac{stbf}, the signal was downsampled to
	32Hz or twice the low-pass frequency 16Hz, resulting in 17 time samples
	between 0.1 and 0.6s. According to the Nyquist Theorem, more samples would not
	contain more information hence the minimum temporal resolution is chosen to reduce
	the dimensionality of the covariance and improve its covariance.
	The activation pattern is the difference between the averages of epochs in response to cued targets and the averages of those in response to non-cued targets.
	We constructed three variations of the spatiotemporal beamformer:
	\ac{stbf} with empirical covariance estimation (\textsc{stbf-emp}) as in
	\cref{sec:empirical_covariance}, \ac{stbf} with
	\ac{loocv}-shrunk covariance estimation (\textsc{stbf-shrunk}) as in
	\cref{sec:shrunk_covariance}, and \ac{stbf} with
	Kronecker-Toeplitz structured covariance estimation (\textsc{stbf-struct} )with \ac{loocv} shrinkage for
	the Kronecker factors as in \cref{sec:structured_covariance}.

@@

	\section{Results}
	\subsection{Minimum required fixed-point iterations}
	The fixed point iteration algorithm described
	in
  \crefrange{eq:fpi_spatial}{eq:fpi-temporal-norm}
  estimates the Kronecker-Toeplitz structured covariance for the
	\textsc{stbf-struct} classifier.
	Fixed-point iteration is an iterative procedure starting from (in our case)
	non-informed initial guesses for the spatial and temporal covariance matrices.
	As a stopping criterion, one could impose a threshold on the difference in
	outcome of successive steps, e.g., based on the covariance norm or the
	classifier accuracy.
	However, few iterations or even just one~\cite{Castaneda2014} suffice to achieve satisfactory performance in practice.

@@

	\subsection{Classifier accuracy for limited training data}
	It is of interest to keep the calibration time before \ac{bci}
	operation as short as possible.
	We mimic this problem by training the classifier with as few training epochs as possible.
	We evaluate the performance of all classifiers for different levels of
	available training data and apply the cross-validation procedure nine times (the number of blocks in the training fold) for all subjects, keeping the
	corresponding number of blocks in the training folds and dropping the rest.
	\cref{fig:accuracy} and \cref{fig:accuracy_ap} show each classifier's
	accuracy relative to the data availability.
	We statistically compare the two newly proposed classifiers,
	\textsc{stbf-struct} and \textsc{stbf-shrunk}, for different levels of training
	data availability using a one-sided paired Wilcoxon rank-sum test with Holm correction for the multiple pairwise comparisons between classifiers.
	We performed this analysis three times: by only using the first trial of a
	block, by averaging epochs across the first two trials of a block and across
	the first five trials of a block.
	Results validated on one trial are reported in \cref{tab:p_values_1}, 2-trial results
	in \cref{tab:p_values_2}, and 5-trial results in \cref{tab:p_values_5}.

  \begin{figure}
		\includegraphics[width=.66\linewidth]{figures/stbf_struct/accuracy.eps}
    \caption[Clasifier accuracy in function of available training data.]{%
      Accuracy of the different classifiers for all 21 subjects relative to the
			number of blocks available for training. One block consists of 135
			epochs and corresponds to 27 seconds of stimulation. Accuracies
			are shown for the evaluation settings averaging over 1, 2, and
			3 trials of testing stimuli.
			\cref{fig:accuracy_ap} contains results for all numbers of trials.
			While \textsc{stbf-emp} is unstable when little training data
			are available,
			regularization of the covariance matrix (\textsc{stbf-shrunk} and
			\textsc{stbf-struct}) drastically improves performance.}
		\label{fig:accuracy}
	\end{figure}
  \begin{table}[p]
		\centering
    \makebox[\textwidth][c]{%
		\input{chapters/stbf_struct_tests_1.tex}
  }
    \caption[Statistical significance of differences in classifier performance (1 repetition).]{%
      Statistical significance of differences in classifier performance.
      $p$-values calculated by one-sided paired Wilcoxon rank-sum test with
			Holm correction using one testing trial for different classifiers
			and levels of data availability. $p$-values $<0.05$ are considered significant
			and marked in grey.}
		\label{tab:p_values_1}
    \bigskip

@@

	When compared to the state-of-the-art \textsc{xdawn+rg} classifier, we conclude
	that \textsc{stbf-struct} reaches similar accuracy when using only one block of
	training data.
	The authors suspect this is due to both
	classifiers having insufficient training information to reach
	satisfactory classification accuracy.
	When more data are available, \textsc{stbf-struct} reaches significantly better accuracies.
	Combined with the benefits laid out in  \cref{sec:parameters_complexity} and
	\cref{sec:interpretability}, this makes it an attractive option for \ac{erp} classification.
	\textsc{stbf-shrunk} does not show decisive accuracy improvements over
	\textsc{xdawn+rg} using a few training blocks, but this improves as the training data increases.

>>>>

  \emph{This chapter was
  published as~\textcite{VanDenKerchove2022}.}

@@

	For the following analyses, we assume that all \ac{eeg} channels are normalized with zero mean and unit variance without loss of generality.
	Solving \cref{eq:minimum_variance} under the linear constraint given by
	\cref{eq:linear_constraint} returns the filter weights $w$ defining the spatiotemporal \ac{lcmv}-beamformer.
	\begin{equation}
    \underset{\mat{w}}{\arg\min}\mat{w}^\intercal \mat{C}
		\mat{w}^\intercal
		\label{eq:minimum_variance}
	\end{equation}
	\begin{equation}
		\mat{a}^\intercal\mat{w} = 1
		\label{eq:linear_constraint}
	\end{equation}
	These weights minimize the variance of the output of the filter while enhancing
	the signal characterized by the constraint.
  $\mat{a} = \text{vec}(\mat{A})$ is the data-driven activation pattern, a template
	of the signal of interest maximizing the difference between two classes of
	epochs, determined as follows:
	\begin{equation}
		\mathbf{a} =
		\frac{1}{|\text{class 1}|}\sum_\text{class 1}\mathbf{x_i} -
		\frac{1}{|\text{class 2}|}\sum_\text{class 2}\mathbf{x_i}
		\label{eq:activation_pattern}
	\end{equation}

	The method of Lagrange multipliers then gives the closed-form solution to the minimization problem posed by
	\cref{eq:minimum_variance} and \cref{eq:linear_constraint} as:
	\begin{equation}
		\mat{w} =
    \frac{\mat{C}^{-1}\mat{a}^\intercal}
		{\mat{a}C^{-1}\mat{a}^\intercal}
		\label{eq:closed_form}
	\end{equation}
	The beamformer can be applied to epochs (unseen or not) as:
	\begin{equation}
		y_i = \mathbf{w}\mathbf{x}_i
		\label{eq:apply_beamformer}
	\end{equation}
	resulting in a scalar output per epoch.
	The linear constraint in \cref{eq:linear_constraint} ensures that the
	beamformer maps epochs containing a target response to a score close to one
	and, conversely, epochs not containing a target response to a score close to
	zero.

	\subsection{Covariance matrix regularization}
	While the spatiotemporal beamformer, in theory, achieves optimal separation
	between target and non-target classes, in analogy to linear discriminant
  analysis~\cite{Treder2016}, it does not always perform well on unseen data.
	The main challenge is to find a good estimator for the inverse covariance
  matrix $\mat{C}^{-1}$ since the real underlying covariance matrix generating the data is, in principle, unknown.

	\subsubsection{Empirical covariance estimation}
	\label{sec:empirical_covariance}
	Earlier spatiotemporal beamformer studies~\cite{Wittevrongel2016,
		Wittevrongel2016a, Wittevrongel2017, Wittevrongel2017a} use the empirical
	covariance and inverse covariance calculated as follows:
	\begin{equation}
    \hat{\mat{C}}_\text{emp} =
		\frac{1}{N-1}\sum^{N}_{i=1}\mat{x}_i\mat{x}_i^\intercal
		\label{eq:empirical_covariance}
	\end{equation}
	\begin{equation}
    \widehat{\mat{C}^{-1}}_\text{emp} = \hat{\mat{C}}_\text{emp}^+
		\label{eq:empirical_inverse_covariance}
	\end{equation}
	The Moore-Penrose pseudoinverse $^+$, ensures a solution exists when
  $\hat{\mat{C}}_\text{emp}$ is singular.
	\cref{fig:kronlda-covs}a and \cref{fig:kronlda-covs}b respectively show examples of the
	empirical estimators of the covariance and the inverse covariance matrices.
	The empirical estimator suffers from performance and
	stability issues if the number of epochs $N$ used or estimation is not much larger than the number of features $CS$~\cite{Stein1956,Khatri1987}.

	\subsubsection{Shrunk covariance estimation}
	\label{sec:shrunk_covariance}
	The shrinkage covariance estimator creates a better-conditioned inversion matrix problem and generally performs better when applied to unseen data.
	The estimators for the covariance and inverse covariance are given by:
	\begin{equation}
    \hat{\mat{C}}_\alpha =
    (1-\alpha) \hat{\mat{C}}_\text{emp}
    + \alpha\frac{\text{Tr}(\hat{\mat{C}}_\text{emp})}{CS}\mathbb{I}
		\label{eq:shrinkage}
	\end{equation}
	\begin{equation}
    \widehat{\mat{C}^{-1}}_\alpha =
    \hat{\mat{C}}^+_\alpha
		\label{eq:shrinkage_inverse}
	\end{equation}
	with $0<\alpha<1$.
	Analogous to $L_2$ regularization of the beamforming problem,
	shrinkage reduces the ratio between the smallest and largest eigenvalues
	of the covariance matrix by strengthening the diagonal.
	\cref{fig:kronlda-covs}c and \cref{fig:kronlda-covs}d respectively show examples of the
	shrunk estimator of the covariance and the inverse covariance matrices.

	Earlier work~\cite{Libert2021} applied shrinkage regularization to \ac{erp}
	decoding with the spatiotemporal beamformer and showed competitive performance
	compared to other state-of-the-art decoding techniques like stepwise LDA or SVM.
	The abovementioned work chooses the shrinkage coefficient $\alpha$ as a fixed hyperparameter.
	However, its optimal value depends on the number of training epochs, the
	covariance matrix's dimensionality, and the independence and variance of the
	data, which can vary across evaluation settings and per session.
	The optimal value for $\alpha$ can be found with a line search using cross-validation, but this can be a costly procedure.
	Methods exist to estimate an optimal shrinkage value from the data directly.
	Most notable among these are the Ledoit-Wolf procedure~\cite{Ledoit2004},
	Rao-Blackwell Ledoit-Wolf~\cite{Chen2010}, and Oracle Approximating Shrinkage~\cite{Chen2010}.
	A more recent estimation method~\cite{Tong2018} emulates a leave-one-out
	cross-validation (\ac{loocv}) scheme expressed by the data-driven closed-form
	estimate:
	\begin{equation}
		\alpha =
		1-\frac{
      \frac{N}{N-1}\text{Tr}(\hat{\mat{C}}_\text{emp}^2)
      - \frac{2}{CS}\left[\text{Tr}(\hat{\mat{C}}_\text{emp})\right]^2
      + \frac{1}{CS}\text{Tr}(\hat{\mat{C}}_\text{emp}^2)
			- \frac{1}{N(N-1)}\sum_{i=1}^N||\mat{x}_i||_2^4
		}
		{
      \frac{N^2 -2N}{(N-1)^2}\text{Tr}(\hat{\mat{C}}_\text{emp}^2)
      - \frac{2}{CS}\left[\text{Tr}(\hat{\mat{C}}_\text{emp})\right]^2
      + \frac{1}{CS}\text{Tr}(\hat{\mat{C}}_\text{emp}^2)
			+ \frac{1}{N(N-1)^2}\sum_{i=1}^N||\mat{x}_i||_2^4
		}
		\label{eq:loocv}
	\end{equation}

@@

	\subsubsection{Spatiotemporal beamforming with Kronecker-Toeplitz structured covariance}
	\label{seq:structured_estimation}
	Exploiting prior knowledge about the spatiotemporal structure of the \ac{eeg} signal leads to a more regularized estimator of the covariance.
	When viewing the example of empirical spatiotemporal \ac{eeg} covariance in
	\cref{fig:kronlda-covs}a, it becomes clear that this matrix consists of a block pattern of repeated, similar matrices.
	Due to the multi-channel nature of the signal, we assume that the covariance of spatiotemporal \ac{eeg} epochs is a Kronecker
	product of two smaller
	matrices~\cite{Munck1992,DeMunck1999,Huizenga2002}, as expressed
	by:
	\begin{equation}
    \hat{\mat{C}}_\text{struct} = \hat{\mat{S}} \otimes \hat{\mat{T}}
		\label{eq:kronecker}
	\end{equation}
	with $\otimes$ the Kronecker product operator.
  $\hat{\mat{S}} \in \mathbb{R}^{C\times C}$ and $\hat{\mat{T}} \in \mathbb{R}^{S\times S}$ respectively correspond to estimators of the spatial and temporal covariance of the data.
	Furthermore, because the temporal covariance of the \ac{eeg}-signal is
	stationary (i.e., it is only dependent on interval length between covarying
	time samples)~\cite{Bijma2003}, it is assumed to have a Toeplitz-matrix structure:
	\begin{equation}
    \hat{\mat{T}}_{i,j} = \hat{\mat{T}}_{i+1,j+1}
		\label{eq:toeplitz}
	\end{equation}
	\Cref{prop:inverse_kronecker} then leads to
	\cref{eq:cov_inverse_kronecker} to estimate the inverse
	covariance.

	\begin{property}
    $(\mat{U} \otimes \mat{V})^+ = \mat{U}^+ \otimes \mat{V}^+$ for any non-singular
    matrices $\mat{U}$ and $\mat{V}$~\cite{Langville2004}.
		\label{prop:inverse_kronecker}
	\end{property}

@@

	Using \cref{eq:closed_form_kronecker} removes the need to calculate the
  full, high dimensional Kronecker product $\hat{\mat{S}}^+\otimes
    \hat{\mat{T}}^+$.
	\Cref{fig:kronlda-covs}e and \cref{fig:kronlda-covs}f respectively show examples of the
	structured covariance and inverse covariance estimators,
	consisting of a spatial Kronecker factor (\cref{fig:kronlda-covs}g and
	\cref{fig:kronlda-covs}h) and a temporal component (\cref{fig:kronlda-covs}i and
	\cref{fig:kronlda-covs}j).

@@

	\begin{subequations}
		\begin{equation}
			\beta_{k+1} =
			1-
			\frac{
        \frac{N}{N-1}\text{Tr}(\tilde{\hat{S}}_{k+1}^2)
        - \frac{2}{C}\left[\text{Tr}(\tilde{\mat{S}}_{k+1})\right]^2
        + \frac{1}{C}\text{Tr}(\tilde{\mat{S}}_{k+1}^2)
				- \frac{1}{N(N-1)}\sum_{i=1}^N
        \left[\text{Tr}(\mat{X}_i\hat{\mat{T}}_k^+\mat{X}_i^\intercal)^2\right]
			}{
        \frac{N^2-2N}{(N-1)^2}\text{Tr}(\tilde{\mat{S}}_{k+1}^2)
        - \frac{2}{C}\left[\text{Tr}(\tilde{\mat{S}}_{k+1})\right]^2
        + \frac{1}{C}\text{Tr}(\tilde{\mat{S}}_{k+1}^2)
				+ \frac{1}{N(N-1)^2}\sum_{i=1}^N
        \left[\text{Tr}(\mat{X}_i\hat{\mat{T}}_k^+\mat{X}_i^\intercal)^2\right]
			}
			\label{eq:spatial_shrinkage}
		\end{equation}
		\begin{equation}
			\gamma_{k+1} =
			1-
			\frac{
        \frac{N}{N-1}\text{Tr}(\tilde{\mat{T}}_{k+1}^2)
        - \frac{2}{S}\left[\text{Tr}(\tilde{\mat{T}}_{k+1})\right]^2
				+ \frac{1}{S}\text{Tr}(\tilde{T}_{k+1}^2)
				- \frac{1}{N(N-1)}\sum_{i=1}^N
        \left[\text{Tr}(\mat{X}_i^\intercal\mat{\mat{S}}_k^+\mat{X}_i)^2\right]
			}{
        \frac{N^2-2N}{(N-1)^2}\text{Tr}(\tilde{\mat{T}}_{k+1}^2)
        - \frac{2}{S}\left[\text{Tr}(\tilde{\mat{T}}_{k+1})\right]^2
        + \frac{1}{S}\text{Tr}(\tilde{\mat{T}}_{k+1}^2)
				+ \frac{1}{N(N-1)^2}\sum_{i=1}^N
        \left[\text{Tr}(\mat{X}_i^\intercal\hat{\mat{S}}_k^+\mat{X}_i)^2\right]
			}
			\label{eq:temporal_shrinkage}
		\end{equation}
	\end{subequations}
	The shrinkage parameters $0<\beta_{k+1}<1$ and $0<\gamma_{k+1}<1$ should be
	re-determined after each iteration.
	The Oracle Approximation Shrinkage method can also be used to determine
	$\beta_{k+1}$ and $\gamma_{k+1}$~\cite{Chen2010,Xie2021} but performs worse for spatiotemporal \ac{eeg} data since not all assumptions are met.

@@

  \subsubsection{\Acl{stbf} classifier}
	Before calculating the \ac{stbf}, the signal was downsampled to
	32Hz or twice the low-pass frequency 16Hz, resulting in 17 time samples
	between 0.1 and 0.6s. According to the Nyquist Theorem, more samples would not
	contain more information hence the minimum temporal resolution is chosen to reduce
	the dimensionality of the covariance and improve its covariance.
	The activation pattern is the difference between the averages of epochs in response to cued targets and the averages of those in response to non-cued targets.
	We constructed three variations of the spatiotemporal beamformer:
	\ac{stbf} with empirical covariance estimation (\textsc{stbf-emp}) as in
	\cref{sec:empirical_covariance}, \ac{stbf} with
	\ac{loocv}-shrunk covariance estimation (\textsc{stbf-shrunk}) as in
	\cref{sec:shrunk_covariance}, and \ac{stbf} with
	Kronecker-Toeplitz structured covariance estimation (\textsc{stbf-struct}) with \ac{loocv} shrinkage for
	the Kronecker factors as in \cref{sec:structured_covariance}.

@@

	\section{Results}
	\subsection{Minimum required fixed-point iterations}
	The fixed point iteration algorithm described
	in
  \crefrange{eq:fpi_spatial}{eq:fpi_temporal_norm}
  estimates the Kronecker-Toeplitz structured covariance for the
	\textsc{stbf-struct} classifier.
	Fixed-point iteration is an iterative procedure starting from (in our case)
	non-informed initial guesses for the spatial and temporal covariance matrices.
	As a stopping criterion, one could impose a threshold on the difference in
	outcome of successive steps, e.g., based on the covariance norm or the
	classifier accuracy.
	However, few iterations or even just one~\cite{Castaneda2014} suffice to achieve satisfactory performance in practice.

@@

	\subsection{Classifier accuracy for limited training data}
	It is of interest to keep the calibration time before \ac{bci}
	operation as short as possible.
	We mimic this problem by training the classifier with as few training epochs as possible.
	We evaluate the performance of all classifiers for different levels of
	available training data and apply the cross-validation procedure nine times (the number of blocks in the training fold) for all subjects, keeping the
	corresponding number of blocks in the training folds and dropping the rest.
	\Cref{fig:accuracy} and \cref{fig:accuracy_ap} show each classifier's
	accuracy relative to the data availability.
	We statistically compare the two newly proposed classifiers,
	\textsc{stbf-struct} and \textsc{stbf-shrunk}, for different levels of training
	data availability using a one-sided paired Wilcoxon rank-sum test with Holm correction for the multiple pairwise comparisons between classifiers.
	We performed this analysis three times: by only using the first trial of a
	block, by averaging epochs across the first two trials of a block and across
	the first five trials of a block.
	Results validated on one trial are reported in \cref{tab:p_values_1}, 2-trial results
	in \cref{tab:p_values_2}, and 5-trial results in \cref{tab:p_values_5}.

  \begin{figure}
		\includegraphics[width=.66\linewidth]{figures/stbf_struct/accuracy.eps}
    \caption[Clasifier accuracy in function of available training data.]{%
      Accuracy of the different classifiers for all 21 subjects relative to the
			number of blocks available for training. One block consists of 135
			epochs and corresponds to 27 seconds of stimulation. Accuracies
			are shown for the evaluation settings averaging over 1, 2, and
			3 trials of testing stimuli.
			\cref{fig:accuracy_ap} contains results for all numbers of trials.
			While \textsc{stbf-emp} is unstable when little training data
			are available,
			regularization of the covariance matrix (\textsc{stbf-shrunk} and
			\textsc{stbf-struct}) drastically improves performance.}
		\label{fig:accuracy}
	\end{figure}

  \fullpagefig{%
		\includegraphics[width=\textwidth]{figures/stbf_struct/accuracy_app.eps}
  }{%
    Accuracy averaging over different numbers of trials.
  }{%
    Accuracy of the different classifiers for all 21 subjects relative to the
		number of blocks available for training. One block consists of 135
		epochs and corresponds to 27 seconds of stimulation. Accuracies
		are shown for the evaluation settings averaging over different
	        numbers of trials, ranging from 1 to 15.

  }{fig:accuracy_ap}

  \begin{table}[p]
		\centering
    \makebox[\textwidth][c]{%
		\input{chapters/stbf_struct_tests_1.tex}
  }
    \caption[Statistical significance of differences in classifier performance (1 repetition).]{%
      Statistical significance of differences in classifier performance.
      $p$-values calculated by one-sided paired Wilcoxon rank-sum test with
			Holm correction using one testing trial for different classifiers
			and levels of data availability. $p$-values $<0.05$ are considered significant
			and marked in grey.}
		\label{tab:p_values_1}
    \bigskip

@@

	When compared to the state-of-the-art \textsc{xdawn+rg} classifier, we conclude
	that \textsc{stbf-struct} reaches similar accuracy when using only one block of
	training data.
	The authors suspect this is due to both
	classifiers having insufficient training information to reach
	satisfactory classification accuracy.
	When more data are available, \textsc{stbf-struct} reaches significantly better accuracies.
	Combined with the benefits laid out in \cref{sec:parameters_complexity}, this
  makes it an attractive option for \ac{erp} classification.
	\textsc{stbf-shrunk} does not show decisive accuracy improvements over
	\textsc{xdawn+rg} using a few training blocks, but this improves as the training data increases.
